---
title: "Radio CovILD, 1-year follow-up"
subtitle: "Supplementary Material"
author: "CovILD study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
always_allow_html: true
output: 
  pdf_document:
    latex_engine: xelatex
bibliography: cov_biblio.bib
csl: elife.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{E\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf", ft_max_row = 6, tab.topcaption = TRUE, ft.align = "center")

set_flextable_defaults(font.size = 11, padding = 2)

use_df_printer()


```

\beginsupplement

\newpage

# Supplementary Figures

```{r fig-ctss-severity, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Predictors of CT abnormality severity at the 1-year follow-up.'}

suppl_figures$ctss_severity$plot

```

__Figure E\@ref(fig:fig-ctss-severity). Predictors of CT abnormality severity at the 1-year follow-up.__

__(A)__ Univariable analysis. Factors associated with CT abnormality severity (classes: 0, 1 - 5, 6 - 10, 11 - 25 CT severity score [CTSS] points) were identified by ordinal logistic regression. Odds ratio (OR) significance was determined by Wald Z test. OR and with 95$\%$ confidence intervals (CI) are presented in a Forest plot. Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.

__(B)__ Multivariable analysis. Independent factors associated with CT abnormality severity (classes: 0, 1 - 5, 6 - 10, 11 - 25 CTSS) were identified by ordinal logistic regression with backward elimination. OR significance was determined by Wald Z test. Model prediction was verified by 20-fold cross-validation (CV), miss-classification error and Cohen's $\kappa$ statistic. OR with 95$\%$ CI are presented in a Forest plot. Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis. Orange: positive correlation, blue: negative correlation, gray: not significant or reference. BMI: body mass index, ref.: reference.

\newpage

```{r fig-ctss-correlation, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Correlation of CTSS and lung opacity.'}

suppl_figures$ctss_correlation$plot

```

__Figure E\@ref(fig:fig-ctss-correlation). Correlation of CTSS and lung opacity.__

Correlation of CTSS and lung opacity at the consecutive follow-ups was investigated with Spearman test. Points represent single observations, blue lines with gray regions represent fitted linear trend with 95$\%$ confidence interval. Correlation coefficients ($\rho$), p values and numbers of complete observations are indicated in the plot captions.